Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-10-01 arxiv/proof_examples.tex
revise

&\exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2} \\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(2)})^\top X^{(2)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right]\\
&+ A^2[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] +\sigma^2(n_1+n_2-p)\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p),

@@

\begin{proof}
\todo{revise the proof}	Using Lemma \ref{largedeviation} again, we can simplify \eqref{revise_eq_val_mtl} as $\val(v)= N_2h(v) \cdot  \left( 1+\OO(p^{-1/2+\e})\right)$, where the function $ h$ is defined as
	%We define the function
	\begin{align}
		h(v) =& \frac{\rho_1}{\rho_2}\left[d^2 +\left( v-1\right)^2\kappa^2\right]\cdot \tr\left[( v^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
		& +  v^2\left[d^2 +\left( v-1\right)^2\kappa^2\right]\cdot \tr\left[( v^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
		& + \left(\frac{\rho_1}{\rho_2} v^2 + 1\right)\sigma^2 \cdot \bigtr{(v^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
	\end{align}
	%\cor Under the setting of Lemma \ref{prop_model_shift_tight},
	Furthermore, the validation loss in \eqref{approxvalid} reduces to
	\be\label{boundv-w}
		g(v)=\left[N_2 h(v) + (N_1+N_2)\sigma^2\right]\cdot \left( 1+\OO(p^{-(1-\e_0)/2+\e})\right)\quad \text{w.h.p.}
	\ee
	for any constant $\e>0$. %\nc Thus for the following discussions, it suffices to focus on the behavior of $h (v)$.
Let $\hat w$ be the minimizer of $h(v)$. Our proof consists of the following two steps.
\squishlist
	\item First, we show that $\hat{w}$ is close to $1$.
	\item Second, with \eqref{boundv-w} we show that $\hat v$ is close to $\hat w$. Then we plug $\hat{v}$ into $\te(\hat{\beta}_2^{\MTL})$ to get \eqref{simple1}.
\squishend
%For the minimizer $\hat w$ of $\val(w)$, we have a similar result as in Proposition \ref{thm_cov_shift}.
For the first step, we will prove the following result.

\begin{proof}
To be consistent with the notation $\hat w$, we change the name of the argument to $w$ in the proof. First it is easy to observe that $h(w)< h(-w)$ for $w> 0$. Second, using Fact \eqref{lem_minv}, it is easy to check that $h(1)=\OO(pd^2 + \sigma^2)$ and $ h(w)\gtrsim p\kappa^2 \gg h(1)$ for $w\ge 2$ or $w\le 1/2$.  Hence it suffices to consider the case $1/2\le w\le 2$.

@@

In sum, we have shown that $w_2\le w\le w_1$. Together with
$$\max\{|w_1 -1|, |w_2 -1|\} =\OO\left(\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2}\right),$$
we conclude \eqref{hatw_add1}.
\end{proof}

%For the rest of this section, we choose parameters that satisfy the following relations: \be\label{choiceofpara}
%pd^2 \sim \sigma^2 \sim 1,\quad p^{-1+c_0} \sigma^2 \le \kappa^2  \le p^{-\e_0-c_0}\sigma^2 ,
%\ee
%for some small constant $c_0>0$. 
Next we prove the following estimate on the optimizer $\hat v$: with high probability,
%\begin{lemma}
%For the isotropic model, we have
\be\label{hatv_add1}
|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
\ee
%\end{lemma}
%\begin{proof}
 In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under \eqref{choiceofpara} we have
$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
Thus we obtain that for $|w-1|\ge 2C\cal E$,
$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
which leads to $g(w) > g(\hat w)$ with high probability by \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with \eqref{hatw_add1}, we conclude \eqref{hatv_add1}.
%\end{proof}
\end{proof}

>>>>

&\exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2} \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(2)})^\top X^{(2)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right] \nonumber\\
&+ A^2[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] +\sigma^2(n_1+n_2-p) \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}

@@

\begin{proof}
We abbreviate
$$h(A):=\exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2} .$$
We first find the minimizer of $h(A)$, denoted by $A^\star$.

%Using Lemma \ref{largedeviation} again, we can simplify \eqref{revise_eq_val_mtl} as $\val(v)= N_2h(v) \cdot  \left( 1+\OO(p^{-1/2+\e})\right)$, where the function $ h$ is defined as
%	%We define the function
%	\begin{align}
%		h(v) =& \frac{\rho_1}{\rho_2}\left[d^2 +\left( v-1\right)^2\kappa^2\right]\cdot \tr\left[( v^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
%		& +  v^2\left[d^2 +\left( v-1\right)^2\kappa^2\right]\cdot \tr\left[( v^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
%		& + \left(\frac{\rho_1}{\rho_2} v^2 + 1\right)\sigma^2 \cdot \bigtr{(v^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
%	\end{align}
%	%\cor Under the setting of Lemma \ref{prop_model_shift_tight},
%	Furthermore, the validation loss in \eqref{approxvalid} reduces to
%	\be\label{boundv-w}
%		g(v)=\left[N_2 h(v) + (N_1+N_2)\sigma^2\right]\cdot \left( 1+\OO(p^{-(1-\e_0)/2+\e})\right)\quad \text{w.h.p.}
%	\ee
%	for any constant $\e>0$. %\nc Thus for the following discussions, it suffices to focus on the behavior of $h (v)$.
%Let $\hat w$ be the minimizer of $h(v)$. Our proof consists of the following two steps.
%\squishlist
%	\item First, we show that $\hat{w}$ is close to $1$.
%	\item Second, with \eqref{boundv-w} we show that $\hat v$ is close to $\hat w$. Then we plug $\hat{v}$ into $\te(\hat{\beta}_2^{\MTL})$ to get \eqref{simple1}.
%\squishend
%%For the minimizer $\hat w$ of $\val(w)$, we have a similar result as in Proposition \ref{thm_cov_shift}.
%For the first step, we will prove the following result.

%To be consistent with the notation $\hat w$, we change the name of the argument to $w$ in the proof. 
First it is easy to observe that $h(w)< h(-w)$ for $w> 0$. Second, using Fact \eqref{lem_minv}, it is easy to check that $h(1)=\OO(pd^2 + \sigma^2)$ and $ h(w)\gtrsim p\kappa^2 \gg h(1)$ for $w\ge 2$ or $w\le 1/2$.  Hence it suffices to consider the case $1/2\le w\le 2$.

@@

In sum, we have shown that $w_2\le w\le w_1$. Together with
$$\max\{|w_1 -1|, |w_2 -1|\} =\OO\left(\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2}\right),$$
we conclude \eqref{hatw_add1}.

%For the rest of this section, we choose parameters that satisfy the following relations: \be\label{choiceofpara}
%pd^2 \sim \sigma^2 \sim 1,\quad p^{-1+c_0} \sigma^2 \le \kappa^2  \le p^{-\e_0-c_0}\sigma^2 ,
%\ee
%for some small constant $c_0>0$. 
Next we prove the following estimate on the optimizer $\hat A$: with high probability,
%\begin{lemma}
%For the isotropic model, we have
\be\label{hatv_add1}
|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
\ee
%\end{lemma}
%\begin{proof}
 In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under \eqref{choiceofpara} we have
$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
Thus we obtain that for $|w-1|\ge 2C\cal E$,
$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
which leads to $g(w) > g(\hat w)$ with high probability by \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with \eqref{hatw_add1}, we conclude \eqref{hatv_add1}.
%\end{proof}
\end{proof}
